---
title: "02-exercises"
author: "Christopher Brown"
date: "April 13, 2016"
output: html_document
---

This exercise uses the **Fuel Economy** data set from the **AppliedPredicitiveModeling** package.

Note: The following will set-up your environment for this exercise. If you get an error stating that the packages have not been found, you need to install those packages.


```{r,echo=FALSE}

<<<<<<< HEAD
packs <-  c('AppliedPredictiveModeling', 'ggplot2', 'magrittr', 'dplyr', 'magrittr')
=======
packs <-  c('AppliedPredictiveModeling', 'ggplot2', 'magrittr', 'dplyr')
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46

for( nm in packs ) { 
  # message(nm)
  if( ! nm  %in% installed.packages()[,1]  ) install.packages(nm)
  library(nm, character.only = TRUE)
}

# Load data set into environment
data(FuelEconomy)
<<<<<<< HEAD
#>?? ... = NULL  # Needed for aesthetics 
=======
... = NULL  # Needed for aesthetics 
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46

```



## Exercise 1

Hint: See `?cars2010`


- After the **Fuel Economy** data is loaded, combine three data sets into one data set. (Note: The name `dat` is very often used in these situations, `data` is a reserved R word.)  

```{r}
<<<<<<< HEAD
dat <- rbind(cars2010, cars2011, cars2012)    # Define 
=======
dat <- bind_rows( cars2010, cars2011, cars2012)    # Define 
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46
```


- What is a good "naive guess" of FE? Show your work

```{r}
naive_guess <- mean(dat$FE)
```


- plot FE (Fuel Econonomy) vs. EngDisp. Plot the naive guess.

```{r}
<<<<<<< HEAD
qplot( y=dat$FE, x=dat$EngDispl ) + geom_hline(yintercept=naive_guess, color="red")
=======
# ... ggplot2  
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46
```

- Sample 10 observations from `dat`
- Plot this data. Add a line for the naive_guess. 


```{r}

set.seed(314) 

# Sample 
<<<<<<< HEAD
samp <- dat %>% sample_n(10)
# to explicitly specify function package also can:
# dplyr::sample_n  

# Plots 
qplot(y = samp$FE, x = samp$EngDispl) + geom_hline(yintercept = naive_guess, color = 'red')
=======
samp <- dat %>% dplyr::sample_n(10) 

# Plots 
...

>>>>>>> 556966077518c2e2067944796437a9e3521a7b46

```


## Exercise 2:


Write a loss functions for calculating:
- Root Mean Square Error
- Mean Absolute Error
- Median Absolute Error

All functions should accept two arguments: 


```{r}

rmse <- function(y,yhat) {
<<<<<<< HEAD
return ((y - yhat)^2 %>% mean %>% sqrt)
}

mae <- function(y, yhat) {
return ((y - yhat) %>% abs %>% mean)
}

medae <- function(y, yhat) { 
return ((y - yhat) %>% median)
=======
  ( y - yhat )^2  %>% mean   %>% sqrt 
}

mae <- function(y, yhat) {
  # ...
}

medae <- function(y, yhat) { 
  # ...
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46
}

```

Use these functions to evaluate the loss/performance of:
- the naive guess

<<<<<<< HEAD
```{r}
y <- samp$FE
yhat <- naive_guess
paste('RMSE naive:', format(round(rmse(y,yhat), 2), nsmall = 2))
paste('MAE naive:', format(round(mae(y,yhat), 2), nsmall = 2))
paste('MEDAE naive:', format(round(medae(y,yhat), 2), nsmall = 2))
```
=======

>>>>>>> 556966077518c2e2067944796437a9e3521a7b46

## Exercise 3: Linear Model and Model Performance 

- Use `lm` to create a linear model fitting the relationship between FE and EngDispl for the cars2010 data set


```{r}
<<<<<<< HEAD
fit.2010 <- lm(FE ~ EngDispl, data = cars2010)
=======
fit.2010 <- ... 
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46
```

- Use your functions to evaluate the training error
- Use your model to:
-- predict the FE for 2011. What is the RMSE errors associated with the predictions.
-- predict the FE for 2012. What is the RMSE errors associated with the predictions.

```{r}

<<<<<<< HEAD
y.2010 <- predict(fit.2010, newdata=cars2010)
y.2011 <- fit.2010 %>% predict(cars2011) 
# newdata=, not data = , otherwise creates preds for 2010 still:
y.2011 <- predict(fit.2010, newdata=cars2011)
y.2012 <- fit.2010 %>% predict(cars2012) 

rmse.2010 <- rmse(cars2010$FE, y.2010)
rmse.2011 <- rmse(cars2011$FE, y.2011)
rmse.2012 <- rmse(cars2012$FE, y.2012)
=======
y.2010 <- ... 
y.2011 <- ... 
y.2012 <- ...

rmse.2010 <- ...
rmse.2011 <- ...
rmse.2012 <- ... 
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46

```


<<<<<<< HEAD
```{r, echo=TRUE}
=======
```{r}
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46
# DO NOT EDIT 
rmse.2010 
rmse.2011
rmse.2012
```


## Exercise 4: 

- Model the fuel economy (`FE`) as a function of `EngDispl`, `NumCyl` and `VarValve` using the `cars2011` data set.  
- Provide betas


```{r}
<<<<<<< HEAD
fit.multivar <- lm(FE ~ EngDispl + NumCyl + VarValveLift, data = cars2011)
summary(fit.multivar)
# kill VarValveLift - insignificant predictor. replace with VarValveTiming:
fit.multivar <- lm(FE ~ EngDispl + NumCyl + VarValveTiming, data = cars2011)
summary(fit.multivar)
betas <- fit.multivar$coefficients
print(betas)
=======
fit.2010 <- ...
betas <- ...
>>>>>>> 556966077518c2e2067944796437a9e3521a7b46
```
